Bringing the Power of eBPF to Open vSwitch

### 2018-10-24 paper.tex
fix afxdp section

AF\_XDP was proposed and upstreamed to Linux kernel since 4.18~\cite{af_xdp_lwn} 
The core idea behind the AF\_XDP is to leverage the XDP eBPF program's
early access to the raw packet, and create a high speed channel from the XDP directly to
a userspace socket interface. In other word, AF\_XDP socket family connects the
XDP packet receving/sending path to the userspace, bypassing the rest of the
Linux networking stacks.
An AF\_XDP socket, called XSK, is create using the normal socket() system
call. Unlike AF\_PACKET which uses the send() and receive() syscalls with packet buffer as parameter,
XSK introduces two rings in userspace: the RX ring and the TX ring.
The userspace program using XSK needs to properly configure and maintain
the RX and TX ring structure in order to receive and send packets.
In addition, all the packet data buffers used in TX/RX ring are allocated from a specific
memory area called UMEM which consists of a number of fixed size chunks.
The UMEM also has two rings: the FILL ring and the COMPLETION ring.
A descriptor in RX/TX ring or in FILL/COMPLETION ring points to the element
in UMEM by its address.  The address is not system's virtual or physical address
but simply an offset within the entire UMEM memory region.

For example, to receive packets, a set of descriptors pointing to empty
packet buffer needs to be filled into the FILL ring.  When packets arrives,
the userspace program checks the RX ring, fetches the packet data from the
desciptors, and refills the empty buffer to the FILL ring structure, in order to receive new
incoming packets.  For sending packets, a set of descriptors pointing the
packet buffer with contents filled to the TX ring then issus send() system call.
It is up to the userspace program to make sure whether the packets have been
sent or not, by checking the COMPLETION ring structure.
In summary, users of XSK needs to properly interact with the following four
rings:
\begin{itemize}
\item FILL ring: for users to fill UMEM addresses to kernel for receiving packets.
\item RX ring: for users to access recevied packets. 
\item TX ring: for users to place packets needed to be sent.
\item COMPLETION ring: for users to make sure packets are sent.
\end{itemize}
Unlike AF\_PACKET which is bound to entire netdev, the XSK is more fine-grained.
XSK is bound to a specific queue id on a device, so only the traffic/flow 
sent to the queue id shows up in the XSK.

@@

\item Create a AF\_XDP socket:
Call socket() syscall to create a XSK,
set up its RX and TX ring buffer, allocate a UMEM region,
and set up FILL/COMPLETION ring of the UMEM.

@@

\item Populate the UMEM FILL ring: Get a couple of UMEM elements and place into FILL ring.
\end{enumerate}
On the other hand, when a AF\_XDP netdev is detached or closed by user,
ovs-vswitchd closes the XSK socket, free the UMEM memory region, and
unload the eBPF program and map.

In order to properly set-up the RX/TX/FILL/COMPLETION rings, we implement a
UMEM memory management API. Basically, it's a list maintaining the unused elements
in UMEM, called UMEM freelist, and function calls to get/put elements into the list.
\subsection{Packet Reception}
\begin{figure}
{\scriptsize
\begin{verbatim}
UMEM: total 8 elements, each has 2K chunk size
                 +-------------------------------+
                 |   |   |   |   |   |   |   |   |
                 +-------------------------------+
            addr:  1   2   3   4   5   6   7  8 
(1) Initial Stat
           -------------------          UMEM freelist
FILL Qu  ...| 1 | 2 | 3 | 4 | ...       {5, 6, 7, 8}
           -------------------
           -------------------
RX ring  ...|   |   |   |   | ...
           -------------------
(2) 4 Packets Arrive
           -------------------          UMEM freelist
FILL Qu  ...|   |   |   |   | ...       {5, 6, 7, 8}
          --------------------
           -------------------
RX ring  ...| 1 | 2 | 3 | 4 | ...
           -------------------
(3) Get 4 free UMEM elemet and refill
           -------------------          UMEM freelist
FILL Qu  ...| 5 | 6 | 7 | 8 | ...       {} 
           -------------------
           -------------------
RX ring  ...| 1 | 2 | 3 | 4 | ...
           -------------------
(4) Pass packets {1, 2, 3, 4} to dpif-netdev
    for parse, lookup, and actions.

(5) Recycle to UMEM freelist, goto step (2)
            +-------------------        UMEM freelist
FILL Qu  ...| 5 | 6 | 7 | 8 | ...       {1, 2, 3, 4}
            +------------------
           -------------------
RX ring  ...|   |   |   |   | ...
           -------------------
\end{verbatim}
}
\vspace{-1.0em}
\caption{An example of how OVS programs the FILL ring and RX ring when
processing incoming AF\_XDP packets.}
\label{afxdprx}
\vspace{-1.0em}
\end{figure}

Figure~\ref{afxdprx} shows how ovs-vswitchd sets up the FILL ring and
the RX ring for receving packets from XSK.  For simplicity to demonstrate
how AF\_XDP works, assume that there are only eight UMEM buffers, with each buffer's
size equals 2KB (?).  Initially at step 1, ovs-vswitchd fills four available UMEM elements into
the FILL ring and waits for incoming packets.  When there are incoming packets,
at step 2, the FILL ring's four buffer elems have been consumed and moved
to the RX ring.  In order to keep receiving packets, ovs-vswitchd gets another
four available UMEM elements from the UMEM freelist, and fills into FILL ring (step 3).
Then ovs-vswitchd creates the metadata needed for the four packet
buffer \{1, 2, 3, 4\}, i.e., struct dp\_packet and struct dp\_packet\_batch, and passes to 
the dpif-netdev layer for parse, lookup and action executions.
Finally, when ovs-vswitchd finishes processing the UMEM buffer, a recycle
mechanism is triggered to place this buffer back to UMEM freelist (step 5).
Step 5 makes sure that there are always available elements in FILL ring, so
that the underlying XDP program in kernel can keep processing packets 
while the userspace ovs-vswitchd is processing the previous
received packets on RX ring.
When step 5 finishes, ovs-vswitchd goes back to step 2, waiting for
new packets.

\subsection{Packet Transmission}
\begin{figure}
{\scriptsize
\begin{verbatim}
 OVS dpif-netdev has 4 packets at UMEM:{1, 2, 3, 4} received
 from eth1 and plan to send to eth2.
(1) Get 4 free elements from eth2's UMEM, copy packet data to it
(2) Create TX descriptors and place into TX ring
           -------------------        eth2 UMEM freelist
TX ring  ...| 4 | 5 | 6 | 7 | ...       {1, 2, 3, 8}
           -------------------
           -------------------
COMPL q  ...|   |   |   |   | ...
           -------------------
(3) Issue sendmsg syscall to eth2's XSK
(4) Packet transmission is completed
           -------------------        eth2 UMEM freelist
TX ring  ...|   |   |   |   | ...       {1, 2, 3, 8}
           -------------------
           -------------------
COMPL q  ...| 4 | 5 | 6 | 7 | ...
           -------------------
(5) Recycle to eth2 UMEM
           -------------------        eth2 UMEM freelist
TX ring  ...|   |   |   |   | ...       {1, 2, 3, 4,
           -------------------           5, 6, 7, 8}
           -------------------
COMPL q  ...|   |   |   |   | ...
           -------------------
(6) Recycle {1, 2, 3, 4} to eth1's UMEM 
\end{verbatim}
}
\vspace{-1.0em}
\caption{An example of how OVS programs the COMPLETION ring and TX ring when
processing incoming AF\_XDP packets from one netdev and sending to another netdev.}
\label{afxdptx}
\vspace{-1.0em}
\end{figure}

@@

Initially, assume ovs-vswitchd receives four packets \{1, 2, 3, 4\} from eth1's XSK
To send to eth2 using eth2's XSK, ovs-vswitchd first gets four packet buffers,
\{4, 5, 6, 7\} from eth2's UMEM freelist and copies the packet data from eth1's UMEM
to eth2's \{4, 5, 6, 7\}.  Then TX descriptors for the four packets are
created and placed into eth2's TX ring (step 2).
At step 3, sendmsg syscall is issued to signal the kernel to start transmit.
As the send in XSK is asynchronos and sendmsg syscall only return return zero when
no error, ovs-vswitch polls the COMPLETION ring to make sure these four
packets have been sent out (step 4).
Once the four packets' descriptors show up at the COMPLETION ring, at step 5,
ovs-vswitchd recycles their UMEM elements back to eth2's UMEM freelist.
In addition, the original four packet buffers from eth1, \{1, 2, 3, 4\}, are
also recycled back to eth1's UMEM freelist.

\subsection{Performance Evaluation}
All of our performance results use a hardware test bed that consists of
two Intel Xeon E5 2440 v2 1.9GHz servers, each with 1 CPU socket and
8 physical cores with hyperthreading enabled.
One server, the target server, has an Intel 40GbE XL710 single port NIC,
and the other server, the source server, has a Netronome NFP-4000 40GbE device.
The target server runs Linux kernel 4.19-rc4, with i40e driver supporting
the AF\_XDP mode.
We installed OVS-AFXDP on the target server, and used the source server to
generates 64-byte single UDP flow packets
at the rate of 19~Mpps using the DPDK library.
On the target server, for demonstrating the best performance,
we disabled the Intel Spectre and Meltdown fixes~\cite{jesper_xdp_perf_drop, retpoline}.

For all our experiments, we fuse a microbenchmark program, called xdpsock,
as the baseline to compare with OVS-AFXDP implementation.
xdpsock is an AF\_XDP sample program doing no packet processing, and can be configured
to do simply dropping all packets, rxdrop, or forwarding the packet to the same port as
it receives, l2fwd.
In our testbed, we measured $19Mpps$ for xdpsock-rxdrop and $17Mpps$ for xdpsock-l2fwd.
For OVS-AFXDP, we conduct similar two experiments, but with OVS's OpenFlow rule installed as
below:
\begin{itemize}
\item OVS-AFXDP rxdrop: Install a single OpenFlow rule to drop every packets,
e.g.,  \texttt{in\_port=eth1, actions=drop}.
\item OVS-AFXDP l2fwd: Install a single OpenFlow rule to forward packet to the same port
as it receives, e.g., \texttt{in\_port=eth1, actions=set\_fields:eth2->in\_port, output:eth1}.
\end{itemize}

Additionally, AF\_XDP provides three operation modes:
XDP\_SKB: works on devices using generic XDP~\cite{genericxdp},
XDP\_DRV: works on devices with XDP support, and
XDP\_DRV ZC (Zero Copy): works on devices with XDP\_XDP zerocopy support.
For all our experiments, we use the XDP\_DRV ZC mode.

\subsubsection{PMD netdev}
Our initial prototype shows pretty terrible performance, only $0.5Mpps$ for rxdrop.
We observe that when dropping/forwarding packets under OVS-AFXDP, two CPUs show up
100\% utilitization:
\begin{itemize}
\item ovs-vswitchd: This is the process keeps doing the send and receive steps
in Figure~\ref{afxdprx} and~\ref{afxdptx}.
\item ksoftirqd/core\_id: This is the kernel software interrupt thread handling the
incoming packets, triggering XDP program to pass packets to the XSK, and also
processing transmission.
\end{itemize}
\mycomment{Compared to xdpsock, ovs-vswitchd adds on top of it a UMEM memory pool,
and packets parsing, lookup and action execution.}
We investigate using Linux perf and our first fix is to enable OVS's PMD-mode,
Poll-Mode-Driver, to the afxdp netdev.

@@

Applying OVS's PMD netdev avoids these problems and improves the rxdrop performance
from $0.5Mpss$ to $3Mpps$.  Ideally, receiving packets from XSK should not need to context
switch into kernel as mentioned in steps~\ref{afxdprx}. So an ideal implementation
should show minimal context switches with most of the CPU time stays in userspace. 
When enabling PMD netdev for afxdp, ovs-vswitch can be
configured to create a {\em dedicated} thread for the packet processing.
In OVS-AFXDP, we create a thread for each XSK's receive queue and the thread keeps polling
the RX ring for new packets.  Each round of receive polling processes a batch of
packets, up to a maximum of 32 packets. In addition, we enable 1GB huge memory page support
to reduce the page fault overhead.  With the above fixes/optimizations, the rxdrop
performance increases to around $10Mpps$

\subsubsection{Memory Management and Locks}\label{memorym}
For every stage of the optimization, we use Linux perf extensively, e.g., perf stat
and perf record/report.  With the above new design, we observe the new bottleneck
is at the UMEM memory pool APIs we introduce to maintain the UMEM freelist, as well as at the
packet metadata allocation, struct dp\_packet, for every received packet.
We introduce two major APIs, umem\_elem\_get(), which gets N free element from the
UMEM freelist, and umem\_elem\_put(), which places back the free UMEM buffer to
the freelist, similar to the concept of producer (put) and consumer (get).
We implemented three different data structures of UMEM memory pool as below
to compare their performance.
\begin{itemize}
\item list\_head: Similar to Linux kernel's list\_head, each element contains its
UMEM element address, and a next pointer points to next free element.
Due to the nature of list\_head, elements are sparse in memory.
Consumer always takes the first element, head, and producer also places back
to the head of the list.
\item ptr\_ring: Similar to Linux kernel's ptr\_ring, an array of elements are
allocated in continuous memory region.
Consumer takes elements from the tail pointer and when placing back, producer, puts
them into where head pointer points to.
\item ptr\_stack: Similar to ptr\_ring, an array of elements are allocated
in continuous memory region, but both consumer and producer take elements
from beginning of the array.
\end{itemize}

Although we allocate one UMEM per netdev, there might be multiple queues per
netdev sharing the same UMEM. As a result, the above three data structures
require a mutex lock before accessing UMEM freelist. Linux perf reports
pthread mutex lock related APIs as ones of the top CPU utilization functions.
We change our design by 1) allocating per-queue UMEM region and 2)
allocating one pmd thread per queue. As a result, no lock is needed because
each queue has only one thread and its own set of UMEM elements.

\subsubsection{Metadata Pre-allocation}
Moving forward, Linux perf shows that the packet metadata allocation
takes a lot of CPU cycles, i.e.,  dp\_packet\_init(), dp\_packet\_set\_data().
So instead of allocating packet metadata at packet reception time, we
implement two data structures and compare their performance:
\begin{itemize}
\item Embedding in packet buffer: Since we already allocate 2K chunk for each
UMEM packet buffer, we reserve the first 256-byte in each UMEM element as struct dp\_packet
and initialize the dp\_packet as much as we can at allocation time.
This is similar to the DPDK mbuf design~\cite{dpdkmbuf}, where a single memory buffer
is used for both packet data and metadata.
\item Separate from packet buffer: This design simply allocates a continuous memory region
storing an array of packet metadata, and initialize them as much as we can.
\end{itemize}

@@

With the above design change, we found that using the ptr\_stack in section~\ref{memorym} with
separate from packet buffer design yields the best performance. Largely due to both data structures
have better spatial locality and more batching friendly, when all data are allocated together.
For example, setting up 32 packet metadatas in an array definitely incurs
less cache misses than setting up in 32 elements in UMEM.
And getting 32 free UMEM elements from ptr\_stack has similar benefits.
With the above decision, the OVS rxdrop shows around $19Mpps$, closed
to the baseline xdpsock rxdrop performance.

>>>>

AF\_XDP was proposed and upstreamed to Linux kernel since 4.18~\cite{af_xdp_lwn} 
The core idea behind the AF\_XDP is to leverage the XDP eBPF program's
early access to the raw packet, and create a high speed channel from the XDP directly to
a userspace socket interface. In other word, AF\_XDP socket family connects the
XDP packet receving/sending path to the userspace, by-passing the rest of the
Linux networking stacks.
An AF\_XDP socket, called XSK, is create using the normal socket() system
call. Unlike AF\_PACKET which uses the send() and receive() syscalls with packet buffer as parameter,
XSK introduces two rings in userspace: the Rx ring and the Tx ring.
The userspace program using XSK needs to properly configure and maintain
the Rx and Tx ring structures in order to receive and send packets.
In addition, all the packet data buffers used in Tx/Rx rings are allocated from a specific
memory area called umem which consists of a number of fixed size chunks.
The umem also has two rings: the Fill ring and the Completion ring.
A descriptor in Rx/Tx ring or in Fill/Completion ring points to the element
in umem by its address. The address is not system's virtual or physical address
but simply an offset within the entire umem memory region.

For example, to receive packets, a set of descriptors pointing to empty
packet buffer needs to be filled into the Fill ring.  When packets arrives,
the userspace program checks the Rx ring, fetches the packet data from the
desciptors, and refills the empty buffer to the Fill ring structure, in order to receive new
incoming packets.  For sending packets, a set of descriptors pointing the
packet buffers with contents filled to the Tx ring, then issus sendmsg() system call.
It is up to the userspace program to make sure whether the packets have been
sent or not, by checking the Completion ring structure.
In summary, users of XSK needs to properly program following four
rings:
\begin{itemize}
\item Fill ring: for users to fill umem addresses to kernel for receiving packets.
\item Rx ring: for users to access recevied packets. 
\item Tx ring: for users to place packets needed to be sent.
\item Completion ring: for users to make sure packets are sent.
\end{itemize}
Unlike AF\_PACKET which is bound to entire netdev, the XSK is more fine-grained.
XSK is bound to a specific queue id on a device, so only the traffic/flow 
sent to the queue id shows up in the XSK.

@@

\item Create a AF\_XDP socket:
Call socket() syscall to create a XSK,
set up its Rx and Tx ring buffer, allocate a umem region,
and set up Fill/Completion ring of the umem.

@@

\item Populate the umem Fill ring: Get a couple of umem elements and place into Fill ring.
\end{enumerate}
On the other hand, when a AF\_XDP netdev is detached or closed by user,
ovs-vswitchd closes the XSK socket, free the umem memory region, and
unload the eBPF program and map.

In order to properly program Rx/Tx/Fill/Completion rings, we implement a
umem memory management layer, call umempool. Basically, it's a data structure
maintaining the unused/available elements in umem, and GET and PUT access APIs.
\subsection{Packet Reception}
\begin{figure}
{\scriptsize
\begin{verbatim}
umem: total 8 elements, each has 2K chunk size
                 +-------------------------------+
                 |   |   |   |   |   |   |   |   |
                 +-------------------------------+
            addr:  1   2   3   4   5   6   7  8 
(1) Initial Stat
           -------------------          umempool
Fill r   ...| 1 | 2 | 3 | 4 | ...       {5, 6, 7, 8}
           -------------------
           -------------------
Rx ring  ...|   |   |   |   | ...
           -------------------
(2) 4 Packets Arrive
           -------------------          umempool
Fill r   ...|   |   |   |   | ...       {5, 6, 7, 8}
          --------------------
           -------------------
Rx ring  ...| 1 | 2 | 3 | 4 | ...
           -------------------
(3) Get 4 free umem elemet and refill
           -------------------          umempool
Fill r   ...| 5 | 6 | 7 | 8 | ...       {} 
           -------------------
           -------------------
Rx ring  ...| 1 | 2 | 3 | 4 | ...
           -------------------
(4) Pass packets {1, 2, 3, 4} to dpif-netdev
    for parse, lookup, and actions.

(5) Recycle to umempool, goto step (2)
            +-------------------        umempool
Fill r  ...| 5 | 6 | 7 | 8 | ...       {1, 2, 3, 4}
            +------------------
           -------------------
Rx ring  ...|   |   |   |   | ...
           -------------------
\end{verbatim}
}
\vspace{-1.0em}
\caption{An example of how OVS programs the Fill ring and Rx ring when
processing incoming AF\_XDP packets.}
\label{afxdprx}
\vspace{-1.0em}
\end{figure}

Figure~\ref{afxdprx} shows how ovs-vswitchd sets up the Fill ring and
the Rx ring for receving packets from XSK.  For simplicity to demonstrate
how AF\_XDP works, assume that there are only eight umem buffers, with each buffer's
size equals 2KB (?).  Initially at step 1, ovs-vswitchd fills four available umem elements into
the Fill ring and waits for incoming packets.  When there are incoming packets,
at step 2, the Fill ring's four buffer elems have been consumed and moved
to the Rx ring.  In order to keep receiving packets, ovs-vswitchd gets another
four available umem elements from the umempool, and fills into Fill ring (step 3).
Then ovs-vswitchd creates the metadata needed for the four packet
buffer \{1, 2, 3, 4\}, i.e., struct dp\_packet and struct dp\_packet\_batch, and passes to 
the dpif-netdev layer for parse, lookup and action executions.
Finally, when ovs-vswitchd finishes processing the umem buffer, a recycle
mechanism is triggered to place this buffer back to umempool (step 5).
Step 5 makes sure that there are always available elements in Fill ring, so
that the underlying XDP program in kernel can keep processing packets 
while the userspace ovs-vswitchd is processing the previous
received packets on Rx ring.
When step 5 finishes, ovs-vswitchd goes back to step 2, waiting for
new packets.

\subsection{Packet Transmission}
\begin{figure}
{\scriptsize
\begin{verbatim}
 OVS dpif-netdev has 4 packets at umem:{1, 2, 3, 4} received
 from eth1 and plan to send to eth2.
(1) Get 4 free elements from eth2's umem, copy packet data to it
(2) Create TX descriptors and place into Tx ring
           -------------------        eth2 umempool
Tx ring  ...| 4 | 5 | 6 | 7 | ...       {1, 2, 3, 8}
           -------------------
           -------------------
Compl r  ...|   |   |   |   | ...
           -------------------
(3) Issue sendmsg syscall to eth2's XSK
(4) Packet transmission is completed
           -------------------        eth2 umempool
Tx ring  ...|   |   |   |   | ...       {1, 2, 3, 8}
           -------------------
           -------------------
Compl r  ...| 4 | 5 | 6 | 7 | ...
           -------------------
(5) Recycle to eth2 umem
           -------------------        eth2 umempool
Tx ring  ...|   |   |   |   | ...       {1, 2, 3, 4,
           -------------------           5, 6, 7, 8}
           -------------------
Compl r  ...|   |   |   |   | ...
           -------------------
(6) Recycle {1, 2, 3, 4} to eth1's umem 
\end{verbatim}
}
\vspace{-1.0em}
\caption{An example of how OVS programs the COMPLETION ring and Tx ring when
processing incoming AF\_XDP packets from one netdev and sending to another netdev.}
\label{afxdptx}
\vspace{-1.0em}
\end{figure}

@@

Initially, assume ovs-vswitchd receives four packets \{1, 2, 3, 4\} from eth1's XSK
To send to eth2 using eth2's XSK, ovs-vswitchd first gets four packet buffers,
\{4, 5, 6, 7\} from eth2's umempool and copies the packet data from eth1's umem
to eth2's \{4, 5, 6, 7\}.  Then TX descriptors for the four packets are
created and placed into eth2's Tx ring (step 2).
At step 3, sendmsg syscall is issued to signal the kernel to start transmit.
As the send in XSK is asynchronos and sendmsg syscall only return return zero when
no error, ovs-vswitch polls the COMPLETION ring to make sure these four
packets have been sent out (step 4).
Once the four packets' descriptors show up at the COMPLETION ring, at step 5,
ovs-vswitchd recycles their umem elements back to eth2's umempool.
In addition, the original four packet buffers from eth1, \{1, 2, 3, 4\}, are
also recycled back to eth1's umempool.

\subsection{Performance Evaluation}
All of our performance results use a hardware test bed consisting of
two Intel Xeon E5 2440 v2 1.9GHz servers, each with 1 CPU socket and
8 physical cores with hyperthreading enabled.
One server, the target server, has an Intel 40GbE XL710 single port NIC,
and the other server, the source server, has a Netronome NFP-4000 40GbE device.
The target server runs Linux kernel 4.19-rc4, with i40e driver supporting
the AF\_XDP zero copy mode~\cite{af_xdp_patch_i40e, af_xdp_patch_zc}.
We installed OVS-AFXDP on the target server, and used the source server to
generates 64-byte single UDP flow packets
at the rate of 19~Mpps using the DPDK library.
On the target server, for demonstrating the best performance,
we disabled the Intel Spectre and Meltdown fixes~\cite{jesper_xdp_perf_drop, retpoline}.

For all our experiments, we use a microbenchmark program, called xdpsock,
as the baseline to compare with OVS-AFXDP implementation.
xdpsock is an AF\_XDP sample program doing no packet processing, and can be configured
to do simply dropping all packets, rxdrop, or forwarding the packet to the same port as
it receives, l2fwd.
In our testbed, we measured $19Mpps$ for xdpsock-rxdrop and $17Mpps$ for xdpsock-l2fwd.
For OVS-AFXDP, we conduct similar two experiments, but with OVS's OpenFlow rule installed as
below:
\begin{itemize}
\item OVS-AFXDP rxdrop: Install a single OpenFlow rule to drop every packets,
e.g.,  \texttt{in\_port=eth1, actions=drop}.
\item OVS-AFXDP l2fwd: Install a single OpenFlow rule to forward packet to the same port
as it receives, e.g., \texttt{in\_port=eth1, actions=set\_fields:eth2->in\_port, output:eth1}.
\end{itemize}

Additionally, AF\_XDP provides three operation modes, from slowest to fastest:
skb mode: works on devices using generic XDP~\cite{genericxdp},
driver mode: works on devices with XDP support, and
zero-copy driver mode: works on devices with XDP\_XDP zerocopy support.
For all our experiments, we use the zero-copy driver mode for showing the best performance.

\subsubsection{PMD netdev}
Our initial prototype shows pretty terrible performance, only $0.5Mpps$ for rxdrop.
We observe that when dropping/forwarding packets under OVS-AFXDP, two CPUs show up
100\% utilitization:
\begin{itemize}
\item ovs-vswitchd: This is the process keeps doing the send and receive steps
in Figure~\ref{afxdprx} and~\ref{afxdptx}.
\item ksoftirqd/core\_id: This is the kernel software interrupt thread handling the
incoming packets, triggering XDP program to pass packets to the XSK, and also
processing transmission.
\end{itemize}
\mycomment{Compared to xdpsock, ovs-vswitchd adds on top of it a umempool,
and packets parsing, lookup and action execution.}
We investigate using Linux perf and our first fix is to enable OVS's PMD-mode,
Poll-Mode-Driver, to the afxdp netdev.

@@

Applying OVS's PMD netdev avoids these problems and improves the rxdrop performance
from $0.5Mpss$ to $3Mpps$.  Ideally, receiving packets from XSK should not need to context
switch into kernel as mentioned in steps~\ref{afxdprx}. So an ideal implementation
should show minimal context switches with most of the CPU time stays in userspace. 
When enabling PMD netdev for afxdp, ovs-vswitch can be
configured to create a {\em dedicated} thread for the packet processing.
In OVS-AFXDP, we create a thread for each XSK's receive queue and the thread keeps polling
the RX ring for new packets.  Each round of receive polling processes a batch of
packets, up to a maximum of 32 packets. In addition, we enable 1GB huge memory page support
to reduce the page fault overhead.

\subsubsection{Memory Management and Locks}\label{memorym}
For every stage of optimizations, we use Linux perf extensively, e.g., perf stat
and perf record/report.  With the above new design, we observe the new bottleneck
is at the umempool APIs we introduce, as well as at the
packet metadata allocation, struct dp\_packet, for every received packet.
We introduce two major APIs, umem\_elem\_get(), which gets N free element from the
umempool, and umem\_elem\_put(), which places back the free umem buffer to
the umempool, similar to the concept of producer (put) and consumer (get).
We implemented three different data structures of umempool as below
to compare their performance.
\begin{itemize}
\item FIFO-list\_head: We reversed first 8-byte of the umem buffer as a next pointer,
pointing to the next available element in umempool, and implement FIFO access get/set
APIs. The list\_head pointer always points to the first available umem element.
\item FILO-ptr\_ring: Similar to Linux kernel's ptr\_ring, an array of elements are
allocated in continuous memory region.
Consumer takes elements from the tail pointer and when placing back, producer, puts
them into where head pointer points to.
\item FIFO-ptr\_array: Similar to Linux kernel's skb\_array~\cite{skbarray},
an array of elements are allocated
in continuous memory region, but both consumer and producer take elements
from beginning of the array.
\end{itemize}

Although we allocate one umem per netdev, there might be multiple queues per
netdev sharing the same umem. As a result, the above three data structures
require a mutex lock before accessing umem freelist. Linux perf reports
pthread mutex lock related APIs as ones of the top CPU utilization functions.
We change our design by 1) allocating per-queue umem region and 2)
allocating one pmd thread per queue. As a result, no lock is needed because
each queue has only one thread and its own set of umem elements.

\subsubsection{Metadata Allocation}
Moving forward, Linux perf shows that the packet metadata allocation
takes a lot of CPU cycles, i.e.,  dp\_packet\_init(), dp\_packet\_set\_data().
So instead of allocating packet metadata at packet reception time, we pre-allocate
the packet metadata and implement two data structures to compare their performances:
\begin{itemize}
\item Embedding in packet buffer: Since we already allocate 2K chunk for each
umem packet buffer, we reserve the first 256-byte in each umem element as struct dp\_packet
and initialize the dp\_packet as much as we can at allocation time.
This is similar to the DPDK mbuf design~\cite{dpdkmbuf}, where a single memory buffer
is used for both packet data and metadata.
\item Separate from packet buffer: This design simply allocates a continuous memory region
storing an array of packet metadata, and initialize them as much as we can.
\end{itemize}

@@

With the above design change, we found that using the FIFO-ptr\_array in section~\ref{memorym} with
separate from packet buffer design yields the best performance. Largely due to both data structures
have better spatial locality and more batching friendly, when all data are allocated together.
For example, setting up 32 packet metadatas in an array definitely incurs
less cache misses than setting up in 32 elements in umem.
And getting 32 free umem elements from ptr\_array has similar benefits.
With the above decision, the OVS rxdrop shows around $19Mpps$, closed
to the baseline xdpsock rxdrop performance.
